---
title: "HAR Final Assessment"
author: "Mathias Barat"
subtitle: Practical Machine Learning Project
job: www.dendropolis.org
logo: null
---

## 1 - SYNOPSIS

### Context

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively.

These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks.

### Problem

One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.

### Goal of the project

Use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

### Dataset information

Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.

Read more: http://groupware.les.inf.puc-rio.br/har#ixzz6ZuaJcLo0

## 2 - DATASETS

### Downloading file:
```{r}

train_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
if (!file.exists("data/training.csv") | !file.exists("data/testing.csv")){
  download.file(train_url, destfile = "data/training.csv")
  download.file(test_url, destfile = "data/testing.csv")
}

training <- read.csv("data/training.csv", na.strings=c("NA","#DIV/0!", ""))
testing <- read.csv("data/testing.csv",na.strings=c("NA","#DIV/0!", ""))
```

### Exploratory Data Analysis
```{r}
summary(training$classe)
unique(training$classe)
training$classe <- factor(training$classe)
```

### Cleaning
Suppress all NA in the dataset
```{r}
training <- training[,colSums(is.na(training)) == 0]
testing <- testing[,colSums(is.na(training)) == 0]
```

### Suppress useless columns
```{r}
# X, user_name, timestamps, windows
training <- training[,-c(1:7)]
testing <- testing[,-c(1:7)]
```

### testing, training and quizz datasets
To avoid errors I rename the datasets and create my own training, testing dataset
```{r}
library(caret)
test_quizz <- testing
inTrain <- createDataPartition(training$classe, p = 0.8, list = FALSE)
temp <- training
training <- temp[inTrain,]
testing <- temp[-inTrain,]
rm(temp)
```

### final dataset
```{r}
# Training dataset:
dim(training)
# Testing dataset:
dim(testing)
# Testing dataset
dim(test_quizz)
```

### Histogram of Y
```{r}
library(ggplot2)
ggplot(training, aes(x=classe)) + geom_histogram(stat = "count", fill="red")+ labs(title="Classe count for the training dataset")
```

## 3 - Recursive Partition modeling

```{r cache=TRUE}
library(ggplot2); library(caret); library(randomForest); library(rpart)

modfit_rpart <- rpart(classe ~ ., data=training, method="class")

predict_rpart <- predict(modfit_rpart, testing, type = "class")

```

```{r}
library(rattle)
fancyRpartPlot(modfit_rpart, cex= 0.5)
```


```{r}

confusionMatrix(predict_rpart, testing$classe)

```
---

## 4 - Random Forest Modeling

```{r}
library(randomForest)
modfit_rforest <- randomForest(classe ~ .,data=training, method = "class", keep.inbag = TRUE)
predict_rforest <- predict(modfit_rforest, testing, type = "class")

modfit_rforest

```
```{r}
varImpPlot(modfit_rforest, cex = 0.8 ) 
```

```{r}
confusionMatrix(predict_rforest, testing$classe)
```
## 5 - Conclusion

### Model choice
The Random forest gets an accuracy of 0.9949 versus a 0.7353. Without any doubt the Random forest is seriously the best for our case.

### Quizz prediction with the test set
```{r}

predict_quizz <- predict(modfit_rforest, test_quizz, type = "class")
predict_quizz

```

